---
title: "Untitled"
author: "ML"
date: "2/1/2022"
output: pdf_document
---

```{r}
library(tidyverse)
```



```{r}
tfs_correlations <-
  read_csv("../data/tfs_correlations_all2.csv")

tfs_correlations <- tfs_correlations |> 
  select(-1)

css_correlations <- read_csv("../data/css_correlations_all2.csv")

css_correlations <- css_correlations |> 
  select(-1)
```


```{r}
tfs_mobility <- read_csv("../data/tfs_means_mobility_withno2.csv")

css_mobility <- read_csv("../data/css_means_mobility_withno2.csv")
```


```{r}
tfs_question_summary_labels <- tfs_correlations |> 
  select(corr_type, Name, Description, Summary, wtdprop, n_colleges)

css_question_summary_labels <- css_correlations |> 
  select(corr_type, Name, Description, Summary, wtdprop, n_colleges)
```


# Working with Google Sheets in R

```{r}
library(googlesheets4)
gs4_deauth() # changes permissions so you don't need to authenticate to read public file
```


Join Google sheet with existing labels
```{r}
css_gsheet <- "https://docs.google.com/spreadsheets/d/1y2ywEeHoe6iLyUVzLaH3Wtxl61rDdzq9xN9DpwQd08M/edit?pli=1#gid=0"
css_groups <- read_sheet(css_gsheet)

css_question_summary2 <- left_join(css_question_summary_labels, css_groups)

css_question_summary_labels <- css_question_summary2
```

```{r}
tfs_gsheet <- "https://docs.google.com/spreadsheets/d/1xaNCVmMfqfrk0MDWJEulUc0yLz35nOgbYgxuzRVPfuA/edit#gid=0"
tfs_groups <- read_sheet(tfs_gsheet)

tfs_question_summary2 <- left_join(tfs_question_summary_labels, tfs_groups)
tfs_question_summary2 <- tfs_question_summary2 |> 
  select(-namethis)

tfs_question_summary_labels <- tfs_question_summary2

tfs_question_summary_labels <- tfs_question_summary_labels |>
  filter(!str_detect(Group_rev, "Z")) 
```

# Make joined files

```{r}
# Drop wtdprop from question summary before saving correlations. We want to keep that column in the summary labels though.

css_question_summary_labels2 <- css_question_summary_labels |> 
  select(-wtdprop)

css_correlations_join <- left_join(css_question_summary_labels2,
                               css_correlations)
```

```{r}
tfs_question_summary_labels2 <- tfs_question_summary_labels |> 
  select(-wtdprop)

tfs_correlations_join <- left_join(tfs_question_summary_labels2,
                               tfs_correlations)
```


## From Notebook 04 01

# TFS Descriptions

## Hanwen, Andy, and Sophia suggested these changes:

```{r}
tfs_df <- tfs_question_summary_labels |>
  mutate(Description = str_remove(Description, "Choose to Attend:"))|>
  mutate(Description = str_remove(Description, "Choice:"))|>
  mutate(Description = str_remove(Description, "Goal:"))|>
  mutate(Description = str_remove(Description, "Future Act:"))|>
  mutate(Description = str_remove(Description, "Reason Attend:"))|>
  mutate(Description = str_remove(Description, "Self Rating:"))|>
  mutate(Description = str_remove(Description, "View:")) |> 
  mutate(Description = str_remove(Description, "[:print:]+:\\s"))
```

## Anything else to fix?

### Removing the colons above worked well. Let's also remove them in the "Your sex:" description

```{r}
tfs_df <- tfs_df |> 
  mutate(Description = ifelse(str_detect(Description, "Your sex:"),
                              "Sex Composition", Description))
```

### Let's add the RACE_ETH variable we created in notebook_02_02.

```{r}
race_variables <- tfs_df |> 
  filter(str_detect(Name, regex("race", ignore_case = TRUE)), # to ignore case
         str_detect(Summary, "yes"))

race_variables <- race_variables |> 
  filter(!str_detect(Name, "GROUP"))

race_variables <- race_variables |> 
  mutate(Label = Description) |> 
  mutate(Name = "RACE_ETHNICITY")

tfs_df <- tfs_df |> 
  filter(!str_detect(Name, "RACE"))

tfs_df <- bind_rows(tfs_df, race_variables)

tfs_df <- tfs_df |> 
  mutate(Description = str_trim(Description, side = c("left")))
```

```{r}
tfs_question_summary_labels <- tfs_df
```


# CSS Descriptions

## Dilay, Evan, and Jaime: Any changes?

## What else to fix?

### Let's remove the "Family Background And Demographics" and "Scales" group:

```{r}
css_df <- css_question_summary_labels

css_df <- css_df |> 
  filter(!str_detect(Group_rev, "Family|Scales"))
```

## Group Suggestions:

```{r}
css_summary2 <- css_df |> 
  mutate(Description = ifelse(Name == "ACT05", # [If this is true do str_replace]
                              str_replace(Description, # where to find the string
                                   "professors", # what string to find
                                   "professor's"), # how to replace match
                              Description))
```

```{r}
css_summary2 <- css_summary2 |> 
  mutate(find_colon = str_count(Description, ":"))
```

The default is to put new variables after the last column. Use `relocate()` to move columns. The default location for `relocate()` is the first column, but you can adjust that with the `.before =` or `.after =` options.

```{r}
css_summary2 <- css_summary2 |> 
  relocate(find_colon, .after = Description)
```

Check a table of the `find_colon` variable to see how bad the problem is.

```{r}
table(css_summary2$find_colon)
```


So we have 15 rows that require additional care. What are our options?


One possibility is to break the `Description` values into two pieces: a string before a colon and a string after (if there is a colon). Then we can use those two pieces to repair the existing values. We can do that with the `separate()` function:

```{r}
css_summary2 <- css_summary2 |> 
  separate(Description, # variable to separate
           c("before_colon", "after_colon"), # the two new columns to create
           sep = "[:]", # where to separate
           remove = FALSE) # don't remove the variable you are separating
```

Our operations above left some whitespace at the start of the strings in the `after_colon` variable. Trim it.

```{r}
css_summary2 <- css_summary2 |> 
  mutate(after_colon = str_trim(after_colon, side = c("left")))
```

Now try to change the existing Description values that have a colon to only take the text after the colon.

```{r}
css_summary2 <- css_summary2 |> 
  mutate(Description = ifelse(find_colon==1, # if this is true...
                              after_colon, # do this...
                              Description)) # if not true do this
```

```{r}
css_summary2 <- css_summary2 |> 
  mutate(Description = ifelse(Name == "ACT05", # [If this is true do str_replace]
                              str_replace(Description, # where to find the string
                                   "professors", # what string to find
                                   "professor's"), # how to replace match
                              Description)) # [If not true do this]
```


```{r}
css_summary2 <- css_summary2 |>
  mutate(before_colon = str_replace_all(before_colon, "Act in College", "Have you")) |>
  mutate(before_colon = str_replace_all(before_colon, "Act in Past Year", "In the past year, have you")) |>
  mutate(before_colon = str_replace_all(before_colon, "Act", "Have you")) |>
  mutate(before_colon = str_replace_all(before_colon, "CSV Performed", "Have you done community/volunteer service")) |>
  mutate(before_colon = str_replace_all(before_colon, "Satisfaction", "How satisfied are you with the")) |>
  mutate(before_colon = str_replace_all(before_colon, "Faculty Provide", "Has the faculty provided")) |>
  mutate(before_colon = str_replace_all(before_colon, "Change", "How have you changed your")) |>
  mutate(before_colon = str_replace_all(before_colon, "Self Rating", "Rate your")) |>
  mutate(before_colon = str_replace_all(before_colon, "View", "Do you agree/disagree that")) |>
  mutate(before_colon = str_replace_all(before_colon, "Success", "Are you")) |>
  mutate(before_colon = str_replace_all(before_colon, "Hours per Week", "How many hours per week did you spend")) |>
  mutate(before_colon = str_replace_all(before_colon, "Goal", "How important to you is")) |>
  mutate(before_colon = str_replace_all(before_colon, "Plans", "Do you plan on"))

```

```{r}
css_summary2 <- css_summary2 |>
  mutate(after_colon = str_to_lower(after_colon))
```

```{r}
css_summary2 <- css_summary2 |>
  mutate(Qmark = "?") |>
  mutate(Description2 = ifelse(find_colon == 0, 
                               before_colon, 
                               paste(before_colon, after_colon, sep = " "))) |>
  mutate(Description2 = ifelse(find_colon == 0, 
                               Description2, 
                               paste(Description2, Qmark, sep = ""))) 
```

```{r}
#css_summary2 <- css_summary2 |> 
  #mutate(Name = str_remove_all(Name, "_9406")) |>
  #mutate(Summary = str_remove_all(Summary, "_9406"))
```


```{r}
css_summary2 <- css_summary2 |> 
  select(corr_type, Name, Description2, Summary, 
         wtdprop, n_colleges, Label_rank, Survey, Survey_Name,
         Label, Group, Group_rev) |> 
  rename(Description = Description2)
```

Remove question mark if rate is in the description.

```{r}
css_summary3 <- css_summary2 |> 
  mutate(Description = ifelse(str_detect(Description, "Rate"),
                              str_remove(Description, "[?]"), Description))
```


```{r}
css_question_summary_labels <- css_summary3
```



## Write all the files

```{r}
write.csv(tfs_question_summary_labels, "../final_project/tfs_question_summary_labels.csv", row.names = FALSE)

write.csv(css_question_summary_labels, "../final_project/css_question_summary_labels.csv", row.names = FALSE)

write.csv(tfs_mobility, "../final_project/tfs_means_mobility_withno.csv", row.names = FALSE)

write.csv(css_mobility, "../final_project/css_means_mobility_withno.csv", row.names = FALSE)

write.csv(tfs_correlations_join, "../final_project/tfs_correlations_join.csv", row.names = FALSE) 
write.csv(css_correlations_join, "../final_project/css_correlations_join.csv", row.names = FALSE) 
write.csv(chetty_fouryr, "../data/chetty_fouryr_shiny.csv", row.names = FALSE)
```