---
title: "Week Three, Class Two"
author: "ML"
date: "1/25/2022"
output: html_document
---

# Load the usual packages
```{r}
library(gganimate)
library(tidyverse)
library(scales)
library(ggthemes)
library(weights)
library(kableExtra)
library(patchwork)
library(DT)
```

# Load the data

## These files have the weighted correlations between every survey question (by response level) and every mobility variable (pooled and by institution type, identified in the `corr_type` variable).

```{r}
tfs_correlations <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_correlations_all.csv")

css_correlations <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_correlations_all.csv")
```

## These files have more detailed labels by question and response level

```{r}
tfs_labels <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels.csv")

css_labels <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_question_summary_labels.csv")
```

# Today's workshop...

Use what you have learned about visualizations and tables to come up with some ways to summarize the correlations.

## For example...

### 10 strongest negative correlations

```{r}
tfs_correlations |> 
  slice_min(k_rank, n = 10) |> 
  select(Description, Summary, k_rank) 
```

### 10 strongest positive correlations
```{r}
css_correlations |> 
  slice_max(k_rank, n = 10) |> 
  select(Description, Summary, k_rank)
```

### Biggest differences in correlations between public and private institutions

```{r}
corr_diff <- tfs_correlations |> 
  filter(corr_type != "Pooled") |> # Filter out the pooled (overall) correlations
  mutate(corr_type = ifelse(str_detect(corr_type, # The "Private, not for profit" label is long. Shorten it.
                                       "Private"), 
                            "Private", corr_type)) |> 
  select(corr_type, Description, Summary, k_rank) |> 
  pivot_wider(names_from = corr_type, # Pivot wider to put correlations in columns
              values_from = k_rank) |> 
  mutate(corr_diff = abs(Public - Private)) |> # New variable to find absolute difference
  slice_max(corr_diff, n = 10)
```

```{r}
corr_diff |> 
  pivot_longer(names_to = "corr_type",
               values_to = "k_rank",
               Private:Public) |> 
  ggplot(aes(x = reorder(Summary, corr_diff),
             y = k_rank,
             color = corr_type)) +
  geom_point() + coord_flip()
```



# Follow up after class

# Still trying to get a summary plot that works. What if we use plotly?

```{r}
# css_labels only has pooled wtdprop. Drop the wtdprop before joining so all rows get labels
css_labels <- css_labels |> 
  select(-wtdprop)

css_join <- left_join(css_correlations, css_labels)
```

## Set up strongest positive and negative correlations by question group

```{r}
css_positive <- css_join |> 
  filter(corr_type == "Pooled",
         Group!="Constructed Scales",
         !str_detect(Summary, "cant")) |> 
  group_by(Group) |> 
  slice_max(k_rank, n = 1) |> # This pulls top 2; could drop to 1 
  select(Group, Description, Summary, k_rank, Label) |> 
  arrange(-k_rank) |> 
  mutate(corr_sign = "Positive",
         abs_k_rank = abs(k_rank))

css_negative <- css_join |> 
  filter(corr_type == "Pooled",
         Group!="Constructed Scales",
         !str_detect(Summary, "cant")) |> 
  group_by(Group) |> 
  slice_min(k_rank, n = 1) |> # This pulls top 2; could drop to 1
  select(Group, Description, Summary, k_rank, Label) |> 
  arrange(-k_rank) |> 
  mutate(corr_sign = "Negative",
         abs_k_rank = abs(k_rank))

css_correlations_sign <- bind_rows(css_positive,
                                   css_negative)

group_levels <- css_correlations_sign |> 
  filter(corr_sign=="Positive") |> 
  arrange(k_rank) |> 
  pull(Group)
  
```

## Make plot combining positive and negative correlations (using absolute value)

```{r}
css_corr_summary_plot <- css_correlations_sign |> 
  mutate(Group = factor(Group,
                        levels = group_levels)) |> 
  ggplot(aes(x = Group, y = abs_k_rank, 
             color = corr_sign,
             text = paste("Group: ", Group, "\n", # text for custom tooltip
                          "Question: ", Description,"\n",
                          "Response: ", Label, "\n",
                          "Correlation: ", round(k_rank,3)))) + 
  geom_point(size = 3) + coord_flip() +
  scale_x_discrete(labels = label_wrap(25)) +
  theme_economist() + 
  guides(color = "none") +
  theme(axis.text.y = element_text(size = 10)) +
  labs(title = "Strongest Positive And Negative Correlations \nIn Each Question Group")

css_corr_summary_plot
```

## Make it interactive with plotly, using tooltips to include question description, level, and corr

See here for setting up font size: https://plotly-r.com/controlling-tooltips.html#styling
```{r}
font <- list(
  family = "Arial",
  size = 14,
  color = "white"
)
label <- list(
  bgcolor = "#232F34",
  bordercolor = "transparent",
  font = font
)

```


```{r}
library(plotly)

ggplotly(css_corr_summary_plot, tooltip = "text") |> 
  layout(legend = list(orientation = "h", x = .25, y = -0.2)) |> 
  style(hoverlabel = label) |> 
  layout(font = font)
```

