---
title: "Week Three, Class Two"
author: "ML"
date: "1/25/2022"
output: html_document
---

# Load the usual packages
```{r}
library(gganimate)
library(tidyverse)
library(scales)
library(ggthemes)
library(weights)
library(kableExtra)
library(patchwork)
library(DT)
```

# Load the data

## These files have the weighted correlations between every survey question (by response level) and every mobility variable (pooled and by institution type, identified in the `corr_type` variable).

```{r}
tfs_correlations <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_correlations_all.csv")

css_correlations <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_correlations_all.csv")
```

## These files have more detailed labels by question and response level

```{r}
tfs_labels <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels.csv")

css_labels <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_question_summary_labels.csv")
```

# Today's workshop...

Use what you have learned about visualizations and tables to come up with some ways to summarize the correlations.


## For example...

### 10 strongest negative correlations

```{r}
css_correlations |> 
  slice_min(k_rank, n = 10) |> 
  select(Description, Summary, k_rank)
```

### 10 strongest positive correlations
```{r}
css_correlations |> 
  slice_max(k_rank, n = 10) |> 
  select(Description, Summary, k_rank)
```

### Biggest differences in correlations between public and private institutions

```{r}
css_correlations |> 
  filter(corr_type != "Pooled") |> # Filter out the pooled (overall) correlations
  mutate(corr_type = ifelse(str_detect(corr_type, # The "Private, not for profit" label is long. Shorten it.
                                       "Private"), 
                            "Private", corr_type)) |> 
  select(corr_type, Description, Summary, k_rank) |> 
  pivot_wider(names_from = corr_type, # Pivot wider to put correlations in columns
              values_from = k_rank) |> 
  mutate(corr_diff = abs(Public - Private)) |> # New variable to find absolute difference
  slice_max(corr_diff, n = 10)
```


