---
title: "Week Three, Class Four"
author: "ML"
date: "1/27/2022"
output: pdf_document
---

# Load the usual packages
```{r}
library(tidyverse)
library(scales)
library(ggthemes)
library(weights)
library(kableExtra)
library(patchwork)
library(DT)
library(plotly)
```

# Load the data

## These files have the weighted correlations between every survey question (by response level) and every mobility variable (pooled and by institution type, identified in the `corr_type` variable).

```{r}
tfs_correlations <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_correlations_all.csv")

css_correlations <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_correlations_all.csv")
```

## These files have more detailed labels by question and response level

```{r}
tfs_labels <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels.csv")

css_labels <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_question_summary_labels.csv")
```

# Collapsing the question groups

Google sheets are here:

- For TFS: https://docs.google.com/spreadsheets/d/1xaNCVmMfqfrk0MDWJEulUc0yLz35nOgbYgxuzRVPfuA/edit#gid=0

- For CSS: https://docs.google.com/spreadsheets/d/1y2ywEeHoe6iLyUVzLaH3Wtxl61rDdzq9xN9DpwQd08M/edit?pli=1#gid=0

- For Mobility: https://docs.google.com/spreadsheets/d/18YH5URpcThSj7QiEnVtLaXD1QDgHQBhIPVb-IzMDoWw/edit?usp=sharing

Edit the Google sheet directly. Make changes to the `Group_rev` column (for TFS and CSS) or the `Description`, `Keep`, and `Group` columns (for mobility) then we'll load those changes into R when you are done.

# Working with Google Sheets in R

```{r}
library(googlesheets4)
gs4_deauth() # changes permissions so you don't need to authenticate to read public file
```

## CSS cleanup

Join Google sheet with existing labels
```{r}
css_gsheet <- "https://docs.google.com/spreadsheets/d/1y2ywEeHoe6iLyUVzLaH3Wtxl61rDdzq9xN9DpwQd08M/edit?pli=1#gid=0"
css_groups <- read_sheet(css_gsheet)

css_labels <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_question_summary_labels.csv")

css_question_summary2 <- left_join(css_labels, css_groups)
```

```{r}
addmargins(table(css_question_summary2$Group, css_question_summary2$Group_rev, exclude = NULL))
```


# TFS

```{r}
tfs_gsheet <- "https://docs.google.com/spreadsheets/d/1xaNCVmMfqfrk0MDWJEulUc0yLz35nOgbYgxuzRVPfuA/edit#gid=0"
tfs_groups <- read_sheet(tfs_gsheet)

tfs_labels <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels.csv")

tfs_question_summary2 <- left_join(tfs_labels, tfs_groups)
tfs_question_summary2 <- tfs_question_summary2 |> 
  select(-namethis)
```

```{r}
addmargins(table(tfs_question_summary2$Group, tfs_question_summary2$Group_rev, exclude = NULL))
```

## Save files to local paths when we are done

### CSS

```{r}
write.csv(css_question_summary2, "../data/css_question_summary_labels.csv", row.names = FALSE)
```

Join new labels to pooled correlations file

```{r}
# Drop wtdprop from question summary before saving correlations. We want to keep that column in the summary labels though.

css_question_summary2 <- css_question_summary2 |> 
  select(-wtdprop)

css_correlations_join <- left_join(css_question_summary2,
                               css_correlations)

write.csv(css_correlations_join, "../data/css_correlations_join.csv",
          row.names = FALSE)
```


### TFS

```{r}
tfs_question_summary2 <- tfs_question_summary2 |>
  filter(!str_detect(Group_rev, "Z")) 

write.csv(tfs_question_summary2, "/Users/lawrence/Desktop/shiny_example/tfs_question_summary_labels.csv", row.names = FALSE)
```

## TFS with new groups

```{r}
# Drop wtdprop from question summary before saving correlations. We want to keep that column in the summary labels though.

tfs_question_summary2 <- tfs_question_summary2 |> 
  select(-wtdprop)

tfs_correlations_join <- left_join(tfs_question_summary2,
                               tfs_correlations)

write.csv(tfs_correlations_join, "/Users/lawrence/Desktop/shiny_example/tfs_correlations_join.csv",
          row.names = FALSE)
```


# Interactive Summary Plot With TFS

```{r}
positive <- tfs_correlations_join |> 
  filter(corr_type == "Pooled", 
         !str_detect(Summary, "cant")) |> 
  select(k_rank, Group, Group_rev, 
         Description, Summary, Label) |> 
  group_by(Group_rev) |>
  slice_max(k_rank, n = 1) |>
  mutate(corr_sign = "Positive") |> 
  arrange(-k_rank)
```

```{r}
negative <- tfs_correlations_join |> 
  filter(corr_type == "Pooled", 
         !str_detect(Summary, "cant")) |> 
  select(k_rank, Group, Group_rev, 
         Description, Summary, Label) |> 
  group_by(Group_rev) |> 
  slice_min(k_rank, n = 1) |> 
  mutate(corr_sign = "Negative") |> 
  arrange(-k_rank)
```

```{r}
correlations_sign <- bind_rows(positive, negative)
```

```{r}
group_levels <- correlations_sign |> 
  filter(corr_sign=="Positive") |> 
  arrange(k_rank) |> 
  pull(Group_rev) 

font <- list(
  family = "Arial",
  size = 14,
  color = "white"
)
label <- list(
  #bgcolor = size,
  bordercolor = "transparent",
  font = font
)
```

```{r}
correlations_summary_plot <- correlations_sign |> 
  mutate(Group_rev = factor(Group_rev,
                            levels = group_levels)) |> 
  ggplot(aes(x = Group_rev, y = k_rank, 
             color = corr_sign,
             text = paste("Group: ", Group_rev, "\n", # text for custom tooltip
                          "Question: ", Description,"\n",
                          "Response: ", Label, "\n",
                          "Correlation: ", round(k_rank,3)))) +
  geom_point(size = 3) + 
  coord_flip() + theme(legend.position = "bottom") +
  scale_color_manual(values = c("red", "forest green"))
```

```{r}
ggplotly(correlations_summary_plot, tooltip = "text") |> 
  layout(legend = list(orientation = "h", x = .25, y = -0.2)) |> 
  style(hoverlabel = label) |> 
  layout(font = font)
```