---
title: "Week One, Class Three"
author: "ML"
date: "1/11/2022"
output: html_document
---

# Morning Recap

```{r}
library(tidyverse)
library(nycflights13)
```

## Example Plots

Top left...
```{r}
flights |> 
  filter(origin %in% c("JFK", "LGA")) |> 
  ggplot(aes(x = origin, y = dep_time, fill = origin)) +
  geom_violin()
```

Top right...
```{r}
flights |> 
  group_by(carrier) |> 
  summarise(mean_distance = mean(distance, na.rm = TRUE)) |> 
  ggplot(aes(x = carrier, y = mean_distance, fill = mean_distance)) +
  geom_col()
```

Bottom left...
```{r}
flights |> 
  group_by(carrier) |> 
  summarise(mean_distance = mean(distance, na.rm = TRUE)) |> 
  ggplot(aes(x = reorder(carrier, -mean_distance), y = mean_distance, 
             fill = mean_distance)) + 
  geom_col()
```

Bottom right...
```{r}
flights |> 
  filter(carrier %in% c("DL", "HA", "OO")) |>
  group_by(carrier, month) |> 
  summarise(mean_dep_delay = mean(dep_delay, na.rm = TRUE)) |> 
  ggplot(aes(x = carrier, y = mean_dep_delay, fill = carrier)) +
  geom_col() + facet_wrap(~month)
```

## Setup

```{r load the mobility data}
mobility <-  read.csv("https://opportunityinsights.org/wp-content/uploads/2018/04/mrc_table2.csv", stringsAsFactors = TRUE)
```

## What are the college tiers?

### Opportunity Insights assigns colleges to selectivity tiers based on the Barron's competitiveness index:

- **Ivy Plus**: Ivy League plus Stanford, MIT, Duke, University of Chicago

- **Other elite schools (public and private)** is Barron's "Most Competitive" category excluding the Ivy Plus schools:
  - ***Most Competitive***: HS rank in top 10% to 20%, HS GPA of A to B+, SAT median between 655 and 800

- **Highly selective** is Barron's "Highly Competitive" tier:
  - ***Highly Competitive***: HS rank in top 20% to 35%, HS GPA of B+ to B, SAT median between 620 and 654

- **Competitive** is Barron's next three tiers:
  - ***Very Competitive***: HS rank in top 35% to 50%, HS GPA no less than B-, SAT median between 573 and 619
  - ***Competitive***: HS Rank in top 50% to 60%, HS GPA of B- to C, SAT median between 500 and 572
  - ***Less Competitive***: HS Rank in top 65%, HS GPA of C, SAT median below 500

- **Noncompetitive** is the remaining two Barron's tiers:
  - ***Noncompetitive***: Generally only require evidence of graduated from HS
  - ***Special***: Professional schools of art, music, nursing, and other specialized disciplines

## Income Ranks By Tier

We have not seen the `k_rank` variable yet. It captures the average rank in the kid's income distribution for students at each college. Use the `summary()` function to get the distribution of this variable.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r summary of k_rank}
summary(mobility$k_rank)
```

These values are across the full range of all 2202 colleges. Let's use `group_by()` with `summarise()` (not `summary()`) to find the mean `k_rank` for each `tier_name`.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r k_rank by tier_name}
mobility |> # This pipe is the new version of %>%
  group_by(tier_name) |> # the variable(s) by which to collect observations
  summarise(mean_rank = mean(k_rank)) # what you want for each group
```

Since each college has a different number of students, we actually want to use the `weighted.mean()` function rather the `mean()` function to find the averages. The `weighted.mean()` function also requires the variable by which to weight the values; in this example, we will use the `count` variable for our weights. 

```{r introducing weighted.mean}
mobility |> 
  group_by(tier_name) |> # the variable(s) by which to collect observations
  summarise(wtd_mean_rank = weighted.mean(k_rank, w = count)) # what you want for each group
```

## Average rank by tier, type, and parent quintile

What would it take to replicate this figure from Opportunity Insights?

https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/notes_slides/1230_figures/mrc_rank_tier.png

We have more tiers than we need for this figure so that doesn't look like the right variable to use. What about `type` or `iclevel`?

Check out the codebook to know what these variables represent:

https://opportunityinsights.org/wp-content/uploads/2018/04/Codebook-MRC-Table-2.pdf

We need the following categories:

- Elite colleges
- Other four year colleges
- Two year colleges
- (We'll deal with the National category later)

So we have to do some data wrangling to make one variable that will have those three groups.

Make a table showing the distribution of the `iclevel` variable.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r}
table(mobility$iclevel)
```


We can start by only keeping the observations from Four-year colleges and Two-year colleges. And let's put all of our work for this figure in a new data frame called `rank_tier_level` based on the existing mobility data frame. 

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###
```{r}
rank_tier_level <- mobility |> 
  filter(iclevel <= 2) # Use the number values since that is how our dataset has them
```

This also works...
```{r}
rank_tier_level <- mobility |> 
  filter(iclevel != 3) # This is still filter, but != means does not equal
```

That dropped 49 observations but there are only 46 colleges in iclevel 3. That's because 3 are NA for iclevel. It's okay to drop them too.

We know we will need to combine some info from the tier_name variable and the iclevel variable. Let's create a new variable in the `rank_tier_level` dataframe called `tier_level` to do so. Recall that the `mutate()` function creates new variables. In this example we are combining it with the `ifelse()` function.

```{r}
rank_tier_level <- rank_tier_level |> 
  mutate(tier_level = ifelse(tier_name %in% c("Ivy Plus", # if this is true...
                                              "Other elite schools (public and private)"),  
                             3, # ... tier_level will take the value of this
                             iclevel)) # ... if the condition is not true, tier_level takes this
```

Make a table with the distribution of this new variable.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

```{r}
table(rank_tier_level$tier_level)
```

We can also use the mutate function to assert that this new variable will be a factor variable and to assign labels.

```{r}
rank_tier_level <- rank_tier_level |> 
  mutate(tier_level = factor(tier_level, # Use tier_level as a factor variable
                             levels = c(3, 1, 2), # reorder so elite is first
                             labels = c("Elite Colleges", # in the level order
                                        "Other Four-Year Colleges",
                                        "Two-Year Colleges")))
```

What is the weighted average `k_rank` in each of these categories?

### REPLACE  THIS LINE WITH YOUR CODE CHUNK ###

```{r}
rank_tier_level |> 
  group_by(tier_level) |> # the variable with your groups
  summarise(wtd_mean_rank =  # what you want for each group
              weighted.mean(k_rank, w = count)) # what you want for each group
```

That gives us the weighted average across *all* parent quintiles. But we want the weighted
average for *each* parent quintile. Are there any other variables that might help us here? And let's save this summarized output in a new data frame called `mean_rank_tier_level` so we can use it later.

```{r}
mean_rank_tier_level <- rank_tier_level |> 
  group_by(tier_level) |> 
  summarise(pq1 = weighted.mean(k_rank_cond_parq1, 
                                              w = count*par_q1),
            pq2 = weighted.mean(k_rank_cond_parq2, 
                                              w = count*par_q2),
            pq3 = weighted.mean(k_rank_cond_parq3, 
                                              w = count*par_q3),
            pq4 = weighted.mean(k_rank_cond_parq4, 
                                              w = count*par_q4),
            pq5 = weighted.mean(k_rank_cond_parq5, 
                                              w = count*par_q5))
```

We have *wide* data (one row for each tier_level) but need *long* data (one row for each tier_level and pq combination). The `pivot_longer` function takes existing variables and collapses them into two new columns: one with the names and one with the values. Then we can use these new columns as the x and y variables in a plot.

```{r pivot_longer example}
mean_rank_tier_level <- mean_rank_tier_level |> 
  pivot_longer(names_to = "pq", # takes existing variable names and puts them in a new column
               values_to = "k_rank", # takes existing values and puts them in a new column
               pq1:pq5) |> # which existing columns to make long
  mutate(k_rank = round(k_rank, 2), # round to two decimal places
         tier_level_replicate = tier_level) # will make labeling easier later
```

Now we have variables we can feed into our aesthetic map.

```{r plot example}
rank_tier_plot <- ggplot(mean_rank_tier_level, aes(x = pq, y = k_rank, 
                             color = tier_level, 
                             group = tier_level_replicate)) + # which points to connect
  geom_point() + 
  geom_line() + # We haven't seen line plots yet
  labs(x = "Parent Quintile", y = "Child Rank",
          title = "Mean Child Rank at Age 34 vs Parent Income Quintile",
       color = "") + # To leave the label of the legend blank
  ylim(c(.3,.8)) # The limits of the y-axis that the figure we are replicating uses 

rank_tier_plot
```

There are many more options for themes in the `ggthemes` package.

```{r}
#install.packages("ggthemes") # hashtag this line after installing
library(ggthemes)
```

I like the simple Tufte theme...

```{r}
rank_tier_plot <- rank_tier_plot + theme_tufte()
rank_tier_plot
```

We don't have annotations yet. But we'll get them later. We also don't have the "National" line. Let's get that by making a new data frame with the overall weighted mean rank by pq. We also want to assign this new variable to the "National" tier_level and make the wide data long again.

```{r}
natl_mean_rank_tier_level <- mobility |> # use mobility this time to include all observations
  summarise(pq1 = weighted.mean(k_rank_cond_parq1, 
                                              w = count*par_q1),
            pq2 = weighted.mean(k_rank_cond_parq2, 
                                              w = count*par_q2),
            pq3 = weighted.mean(k_rank_cond_parq3, 
                                              w = count*par_q3),
            pq4 = weighted.mean(k_rank_cond_parq4, 
                                              w = count*par_q4),
            pq5 = weighted.mean(k_rank_cond_parq5, 
                                              w = count*par_q5)) |> 
  pivot_longer(names_to = "pq", # takes existing variable names and puts them in a new column
               values_to = "k_rank", # takes existing values and puts them in a new column
               pq1:pq5) |> # which existing columns to make long
  mutate(k_rank = round(k_rank, 2), # round to two decimal places
         tier_level = "National",
         tier_level_replicate = tier_level) # will make labeling easier later
```

Let's add the points for these averages to our existing plot.

```{r}
rank_tier_plot2 <- rank_tier_plot + #everything from our original plot
  geom_point(data = natl_mean_rank_tier_level, # need to include data and aes for new df
             aes(x = pq, y = k_rank, 
                 color = tier_level)) +
  geom_line(data = natl_mean_rank_tier_level, # need to include data and aes for new df
            aes(x = pq, y = k_rank, 
                color = tier_level),
            linetype = "longdash") # so we can see the natl line differently

rank_tier_plot2
```

Lots of annotations would make this messy. Let's finish by making this plot interactive instead. We'll use the `plotly` package for that. Click "No" if R asks you to update loaded packages.

```{r install and load plotly}
#install.packages("plotly") # hashtag this line after installation
library(plotly)
```

All you have to do to make the plot interactive is wrap it in the `ggplotly()` function and add the tooltip() option (if you want to customize the tooltip).

```{r plotly example}
ggplotly(
  rank_tier_plot2,
  tooltip = c("pq","k_rank","tier_level")
)
```
