a---
title: "Week Four, Class One"
author: "ML"
date: "1/31/2022"
output: pdf_document
---

# Load the usual packages
```{r}
library(tidyverse)
library(scales)
library(ggthemes)
library(weights)
library(kableExtra)
library(patchwork)
library(DT)
library(plotly)
```

# Load the data

# CLEAN UP BEFORE CLASS!

## These files add the weighted correlations with the two constructed mobility variables (k_rank_cond_parq123 and kq345_cond_parq1).

```{r}
tfs_correlations <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_correlations_all2.csv")

css_correlations <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_correlations_all2.csv")
```

## These files have more detailed labels by question and response level

```{r}
tfs_labels <-
  read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels.csv")

css_labels <- read_csv("../data/css_question_summary_labels.csv")
```


# Collapsing the question groups

Google sheets are here:

- For TFS: https://docs.google.com/spreadsheets/d/1xaNCVmMfqfrk0MDWJEulUc0yLz35nOgbYgxuzRVPfuA/edit#gid=0

- For CSS: https://docs.google.com/spreadsheets/d/1y2ywEeHoe6iLyUVzLaH3Wtxl61rDdzq9xN9DpwQd08M/edit?pli=1#gid=0

- For Mobility: https://docs.google.com/spreadsheets/d/18YH5URpcThSj7QiEnVtLaXD1QDgHQBhIPVb-IzMDoWw/edit?usp=sharing

Edit the Google sheet directly. Make changes to the `Group_rev` column (for TFS and CSS) or the `Description`, `Keep`, and `Group` columns (for mobility) then we'll load those changes into R when you are done.

# Working with Google Sheets in R

```{r}
library(googlesheets4)
gs4_deauth() # changes permissions so you don't need to authenticate to read public file
```

## CSS cleanup

Join Google sheet with existing labels
```{r}
css_gsheet <- "https://docs.google.com/spreadsheets/d/1y2ywEeHoe6iLyUVzLaH3Wtxl61rDdzq9xN9DpwQd08M/edit?pli=1#gid=0"
css_groups <- read_sheet(css_gsheet)

css_question_summary2 <- left_join(css_labels, css_groups)
```


# TFS

```{r}
tfs_gsheet <- "https://docs.google.com/spreadsheets/d/1xaNCVmMfqfrk0MDWJEulUc0yLz35nOgbYgxuzRVPfuA/edit#gid=0"
tfs_groups <- read_sheet(tfs_gsheet)

tfs_question_summary2 <- left_join(tfs_labels, tfs_groups)
tfs_question_summary2 <- tfs_question_summary2 |> 
  select(-namethis)
```

```{r}
# These include the constructed mobility variables:
# k_rank_cond_parq123 and kq345_cond_parq1

tfs_mobility <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_means_mobility_withno2.csv")

write.csv(tfs_mobility, "../data/tfs_means_mobility_withno_shiny.csv",
          row.names = FALSE)

css_mobility <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_means_mobility_withno2.csv")

write.csv(css_mobility, "../data/css_means_mobility_withno_shiny.csv",
          row.names = FALSE)

# No changes here
chetty_fouryr <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/chetty_fouryr.csv")

write.csv(css_mobility, "../data/chetty_fouryr_shiny.csv",
          row.names = FALSE)

```
# Before final cleanup, here are the files our app uses:

```{r}

# These include revised groups
tfs_df <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_question_summary_labels_shiny.csv")

css_df <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_question_summary_labels_shiny.csv")

# These include the constructed mobility variables:
# k_rank_cond_parq123 and kq345_cond_parq1

tfs_mobility <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_means_mobility_withno_shiny.csv")

css_mobility <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_means_mobility_withno_shiny.csv")

# These include revised groups
tfs_correlations_join <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_correlations_join_shiny.csv")

css_correlations_join <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_correlations_join_shiny.csv")

# No changes here
chetty_fouryr <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/chetty_fouryr_shiny.csv")
```

Let's make some final changes before writing these files to the app folder

# TFS Descriptions

## Hanwen, Andy, and Sophia suggested these changes:

```{r}
tfs_df <- tfs_df |>
  mutate(Description = str_remove(Description, "Choose to Attend:"))|>
  mutate(Description = str_remove(Description, "Choice:"))|>
  mutate(Description = str_remove(Description, "Goal:"))|>
  mutate(Description = str_remove(Description, "Future Act:"))|>
  mutate(Description = str_remove(Description, "Reason Attend:"))|>
  mutate(Description = str_remove(Description, "Self Rating:"))|>
  mutate(Description = str_remove(Description, "View:")) |> 
  mutate(Description = str_remove(Description, "[:print:]+:\\s"))
```

## Anything else to fix?

### Removing the colons above worked well. Let's also remove them in the "Your sex:" description

```{r}
tfs_df <- tfs_df |> 
  mutate(Description = ifelse(str_detect(Description, "Your sex:"),
                              "Sex Composition", Description))
```

### Let's add the RACE_ETH variable we created in notebook_02_02.

```{r}
race_variables <- tfs_df |> 
  filter(str_detect(Name, regex("race", ignore_case = TRUE)), # to ignore case
         str_detect(Summary, "yes"))

race_variables <- race_variables |> 
  filter(!str_detect(Name, "GROUP"))

race_variables <- race_variables |> 
  mutate(Label = Description) |> 
  mutate(Name = "RACE_ETHNICITY")

tfs_df <- tfs_df |> 
  filter(!str_detect(Name, "RACE"))

tfs_df <- bind_rows(tfs_df, race_variables)
```

# CSS Descriptions

## Dilay, Evan, and Jaime suggested these changes:

## Anything else to fix?

### Let's remove the "Family Background And Demographics" and "Scales" group:

```{r}
css_df <- css_df |> 
  filter(!str_detect(Group_rev, "Family|Scales"))
```


# Write all these changes to csv files in your app folder

```{r}
write.csv(tfs_df, "tfs_question_summary_labels.csv", row.names = FALSE)
write.csv(css_df, "css_question_summary_labels.csv", row.names = FALSE)

write.csv(tfs_mobility, "tfs_means_mobility_withno.csv", row.names = FALSE)

write.csv(css_mobility, "css_means_mobility_withno.csv", row.names = FALSE)

write.csv(tfs_correlations_join, "tfs_correlations_join.csv", row.names = FALSE) 
write.csv(css_correlations_join, "css_correlations_join.csv", row.names = FALSE) 

write.csv(chetty_fouryr, "chetty_fouryr.csv", row.names = FALSE)
```

# The App!

Install a couple more packages for everything to run properly:

```{r}
install.packages(c("shinyfullscreen", "shinythemes"))
```


The most up to date version is here:

https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/final_project/app.R
