# Getting Started 

Load the packages and data

```{r, warning = FALSE, error = FALSE, message = FALSE}
library(tidyverse)
library(weights)
library(scales)
```

# Quick Morning Check-in

Consult codebooks or data documentation files : https://www.gapminder.org/data/

Thinking about when and why to use logs of income

```{r}
install.packages("gapminder") # hashtag this line after installing
library(gapminder)

# example without logs

gapminder |> 
  filter(year == 2007) |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  scale_x_continuous(labels = dollar_format())
```

```{r}
# example with logs

gapminder |> 
  filter(year == 2007) |> 
  ggplot(aes(x = gdpPercap, y = lifeExp)) +
  geom_point() +
  coord_trans(x = "log10") +
  scale_x_continuous(labels = dollar_format())
```

Do logs matter for our data?

```{r}
mrc <- read_csv("https://opportunityinsights.org/wp-content/uploads/2018/04/mrc_table2.csv")

# Example without logs
mrc |> 
  ggplot(aes(x = par_median)) +
  geom_density() +
  geom_vline(xintercept = 
               mrc$par_median[mrc$name=="Middlebury College"],
             color = "blue") +
  geom_vline(xintercept = 
               mean(mrc$par_median, na.rm = TRUE),
             color = "forest green") +
  geom_vline(xintercept = 
               median(mrc$par_median, na.rm = TRUE),
             color = "red") +
  labs(title = "Median Income By College",
       subtitle = "Red = Median, Green = Mean, Blue = Middlebury") +
  scale_x_continuous(labels = dollar_format())
```

```{r}
# Example with logs
mrc |> 
  ggplot(aes(x = par_median)) +
  geom_density() +
  geom_vline(xintercept = 
               mrc$par_median[mrc$name=="Middlebury College"],
             color = "blue") +
  geom_vline(xintercept = 
               mean(mrc$par_median, na.rm = TRUE),
             color = "forest green") +
  geom_vline(xintercept = 
               median(mrc$par_median, na.rm = TRUE),
             color = "red") +
  labs(title = "Logged Median Income By College",
       subtitle = "Red = Median, Green = Mean, Blue = Middlebury") +
  scale_x_continuous(labels = dollar_format()) +
  coord_trans(x = "log10")
```

# Returning to our correlations and associations

```{r}
css_means <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/css_means_mobility_withno.csv")

tfs_means <- read_csv("https://raw.githubusercontent.com/mjclawrence/soci1230_w22/main/data/tfs_means_mobility_withno.csv")
```

Let's remove the `_mean` at the end of the columns names in both dataframes.

```{r}
css_means <- css_means |> 
  pivot_longer(names_to = "variable",
               values_to = "value",
               ends_with("_mean")) |> 
  mutate(variable = str_remove(variable, "_mean")) |> 
  pivot_wider(names_from = "variable",
              values_from = "value")
```

```{r}
tfs_means <- tfs_means |> 
  pivot_longer(names_to = "variable",
               values_to = "value",
               ends_with("_mean")) |> # A stringr-like function for columns
  mutate(variable = str_remove(variable, "_mean")) |> 
  pivot_wider(names_from = "variable",
              values_from = "value")
```

# Continuing correlations

We ended class yesterday with a task: find the association between the CSS variable `HPW10.hours0` (zero hours spent partying per week) and `k_rank_cond_parq5` (average adult income rank for kids who grow up in the top quintile). What did you expect? What did you find?

It's a good idea to break this into pieces so you can use parts in later analyses. First save the "final" data frame you will need for any summaries or plots. Get the question variables and create the weights.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

Then you can use that dataframe to summarise the correlation.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

And use the same dataframe to create the scatterplots.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

Any other variables and specific responses you want to look at?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

The bonus question at the end of yesterday's notebook asked if there is a way to easily work with *all* the responses to a question at once. What do you think?

Let's try `SLFCHG30` from the CSS (self perception of change in religious beliefs and convictions during college) and `k_rank`. What is a hypothesis for the overall associations? Would you expect the associations to vary by college type?

Getting the variables and creating the weight is the same as we have been doing, but we'll need to add `type` (public and private). The pivot_longer is slightly different because we can't use the shortcut in the HPW variables to find columns that contain "hours".

```{r}
cor_df <- css_means |> 
  select(campus_id, 
         n_responses, 
         type,
         k_rank,
         starts_with("SLFCHG30")) |> 
  mutate(n_responses_nona = n_responses * (1-SLFCHG30.propna)) |> 
  pivot_longer(names_to = "response_level",
               values_to = "prop",
               (starts_with("SLFCHG30") & !contains("propna"))) |>
  mutate(response_level = str_remove(response_level,
                                     "SLFCHG30.")) |> 
  mutate(response_level = factor(response_level,
                                 levels = c("muchweaker",
                                            "weaker",
                                            "nochange",
                                            "stronger",
                                            "muchstronger"),
                                 labels = c("Much weaker",
                                            "Weaker",
                                            "No change",
                                            "Stronger",
                                            "Much stronger")))
```

The new part is that we want to use `group_by()` to get a separate correlation for each response level. Try it.

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

Can we get this by type of college?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

Might be easier to understand what's going on if we only have two rows (the types) with the response levels in columns and the weighted correlations in the cells. How do we do that?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###


How could we get scatterplots for all the response levels to this question?

### REPLACE THIS LINE WITH YOUR CODE CHUNK ###

