---
title: "setup_css"
author: "ML"
date: "1/4/2022"
output: html_document
---

# Load packages and set wd
```{r}
library(tidyverse)
library(weights)
```

# Load data

The `collnames` file links campus_id, ACERECODE, and super_opeid.

```{r}
collnames <- read_csv("../data/collnames.csv")
```

The `college_mobility` file has chetty variables and css variables and tfs matched variables

```{r}
college_mobility <- read_csv("../data/college_mobility.csv")
```

The `chetty_table2.csv` file has all the mobility variables

```{r}
chetty <- read_csv("../data/chetty_table2.csv")
chetty_vars <- chetty |> 
  select(super_opeid, mr_kq5_pq1,
         mr_ktop1_pq1,
                 k_rank, k_q5, k_top10pc,
                 k_top5pc, k_top1pc,
                 starts_with("k_rank_cond_parq"),
                 starts_with("kq5_cond_parq"),
                 par_q1, par_q5)

chetty_vars <- left_join(collnames, chetty_vars)

chetty_vars <- chetty_vars |> 
  select(-c(ACERECODE, name, super_opeid))
```

The `css_questions_usable` file notes which questions are used in later analyses. It also has the question descriptions.

```{r}
css_names.labeled <- read_csv("../data/css_questions_usable.csv")
```

Keep the questions we are using.

```{r}
css_names.labeled <- css_names.labeled %>% 
  filter(usable_m == 1) |> 
  select(Name, Description, n_asked, usable_m)
```

## Prepare data for analyses

Arrange the variable names and descriptions alphabetically

```{r}
keep <- css_names.labeled %>% 
  dplyr::select(1:3) %>%
  arrange("Name")
```

Filter mobility data for column names we care about 

```{r}
usable_questions <- css_names.labeled$Name
all_colnames <- colnames(college_mobility)
intersection <- Reduce(intersect,list(usable_questions,all_colnames))
#mobility columns here
#mobility_cols <- colnames(college_mobility)[0:65]
#add campus_id
mobility_cols <- colnames(college_mobility)[1]
all_filter_cols<- c(intersection,mobility_cols)
#filter for desired columns
#college_mobility.filtered <- college_mobility[, all_filter_cols]

college_mobility.filtered <- college_mobility |> 
  select(all_of(all_filter_cols))
```

Check the NAs

```{r}
colSums(is.na(college_mobility.filtered))
```

Only keep colleges with more than 100 responses

```{r}
x <- college_mobility.filtered %>% 
  count(campus_id)

x <- x  %>% filter(n>100)
college_mobility.filtered <- left_join(x, college_mobility.filtered, by = "campus_id")
```

Get a vector of campus_id values for colleges we are using

```{r}
campus_id <- c(unique(college_mobility.filtered[,"campus_id"]))
```

## Find proportion missing for each question by school

```{r}
set.seed(1)

college_proportion_answered <- list()
test <- list()

#loop through every campusid
for(i in unique(college_mobility.filtered$campus_id)){
  #get the 
  cm<-college_mobility.filtered %>% 
    filter(campus_id == i) 
  
  na<-colSums(is.na(cm))
  not.na<-colSums(!is.na(cm))
  prop<- na/(na+not.na)
  test[[i]]<- prop
  
}

# Bind rows from each iteration
big_data = do.call(rbind, test)

# Make matrix a data frame
big_data <- as.data.frame(big_data)

# Drop prop missing campus_id
big_data <- dplyr::select(big_data, -c("campus_id"))

# Bind vector and data frame
big_data <- cbind.data.frame(campus_id, big_data)
```

## Calculate means of all variables by college

```{r}
css_means <- college_mobility.filtered |> 
  group_by(campus_id) |> 
  summarise(
            across(starts_with("ACT"),
                   list(mean = ~mean(.x==3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()), 
                   .names = "{.col}.frequently_{.fn}"),
            across(starts_with("COLACT02"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.race/culture_workshop_{.fn}"),
            across(starts_with("COLACT07"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.honors_classes_{.fn}"),
            across(starts_with("COLACT11"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.pt_job_off_{.fn}"),
            across(starts_with("COLACT12"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.pt_job_on_{.fn}"),
            across(starts_with("COLACT13"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.roommate_dif_eth/race_{.fn}"),
            across(starts_with("COLACT14"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.greeklife_member_{.fn}"),
            across(starts_with("COLACT16"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.study_abroad_{.fn}"),
            across(starts_with("COLACT17"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.in_eth/race_org_{.fn}"),
            across(starts_with("COLACT18"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.internship_{.fn}"),
            across(starts_with("COLACT19"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.campus_protest_{.fn}"),
            across(starts_with("COLACT21"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.leadership_training_{.fn}"),
            across(starts_with("COLACT23"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.student_gov_{.fn}"),
            across(starts_with("COLACT27"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.took_remedial_course_{.fn}"),
            across(starts_with("COLACT32"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.community_col_transfer_{.fn}"),
            across(starts_with("COLACT33"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.4y_col_transfer_{.fn}"),
            across(starts_with("COLACT34"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.college_transfer_{.fn}"),
            across(starts_with("COLACT36"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.voted_state/nat_election_{.fn}"),
            across(starts_with("COLACT40"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.ft_job_{.fn}"),
            across(starts_with("COLACT41"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.research_w/_prof_{.fn}"),
            across(starts_with("CSV"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.{.fn}"),
            across(starts_with("FACPRV"),
                   list(mean = ~mean(.x==3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.frequently_{.fn}"),
            across(starts_with("GENACT"),
                   list(mean = ~mean(.x==3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.frequently_{.fn}"),
            across(starts_with("GOAL"),
                   list(mean = ~mean(.x>=3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.important_{.fn}"),
            across(starts_with("HPW"),
                   list(mean = ~mean(.x>=6, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.11hrs_plus_{.fn}"),
            across(starts_with("MAJOR"),
                   list(mean = ~mean(.x==3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.business_major_{.fn}"),
            across(starts_with("NATENG"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.{.fn}"),
            across(starts_with("PLANS"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.yes_{.fn}"),
            across(starts_with("POLIVIEW"),
                   list(mean = ~mean(.x>=4, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.liberal/farleft_{.fn}"),
            across(starts_with("RACE"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.{.fn}"),
            across(starts_with("RATE"),
                   list(mean = ~mean(.x>=5, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.abv_avg_{.fn}"),
            across(starts_with("REENROLL"),
                   list(mean = ~mean(.x>=4, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.def_yes_{.fn}"),
            across(starts_with("SREL"),
                   list(jewish_mean = ~mean(.x==6, na.rm = TRUE),
                        catholic_mean = ~mean(.x==12, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.{.fn}"),
            across(starts_with("SATIS"),
                   list(mean = ~mean(.x>=4, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.satisfied_or_very_satisfied_{.fn}"),
            across(starts_with("SEX"),
                   list(mean = ~mean(.x==2, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.female_{.fn}"),
            across(starts_with("SLFCH"),
                   list(mean = ~mean(.x>=4, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.stronger_or_much_stronger_{.fn}"),
            across(starts_with("SUCC"),
                   list(mean = ~mean(.x==3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.{.fn}"),
            across(starts_with("VIEW"),
                   list(mean = ~mean(.x>=3, na.rm = TRUE), 
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.agree_somewhat_or_strongly_{.fn}"),
            across(starts_with("ACADEMIC_SELFCONCEPT"),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.scale_{.fn}"),
            across(starts_with("SOCIAL_SELFCONCEPT"),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.scale_{.fn}"),
            across(starts_with("SOCIAL_AGENCY"),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.scale_{.fn}"),
            across(starts_with("FAC_INTERACTION"),
                   list(mean = ~mean(.x, na.rm = TRUE),
                        propna = ~sum(is.na(.)) / n()),
                   .names = "{.col}.scale_{.fn}"),
            n_responses = n()
  ) %>%
  select(sort(names(.))) %>%
  relocate(campus_id, n_responses) %>%
  mutate(across(everything(), ~na_if(.,"NaN")))
```

Keep only the variable names in a vector

```{r}
varnames_original <- gsub("(.+?)(\\..*)", "\\1", names(css_means))
varnames_summarize <- names(css_means)
varnames <- bind_cols(varnames_original, varnames_summarize)
colnames(varnames) <- c("Name", "Summarized")
varnames <- left_join(varnames, css_names.labeled)
varnames <- varnames |> 
  filter(!str_detect(Summarized, "propna"))
```


## Correlations

Merge the variable means by college with the mobility variables

```{r}
css_means_mobility <- left_join(css_means, chetty_vars)
```

Calculate correlations between all college means and all mobility variables

```{r}
css_correlations <- css_means_mobility |> 
  select(-c(campus_id, ends_with("propna"))) |> 
  summarise_at(vars(-n_responses), 
                   list(cor_mr_kq5_pq1 = ~wtd.cor(.x, mr_kq5_pq1, 
                                                  weight = n_responses)[1],
                        cor_mr_ktop1_pq1 = ~wtd.cor(.x, mr_ktop1_pq1, 
                                                    weight = n_responses)[1],
                        cor_k_rank = ~wtd.cor(.x, k_rank, 
                                              weight = n_responses)[1],
                        cor_k_q5 = ~wtd.cor(.x, k_q5, 
                                            weight = n_responses)[1],
                        cor_k_top10pc = ~wtd.cor(.x, k_top10pc, 
                                                 weight = n_responses)[1],
                        cor_k_top5pc = ~wtd.cor(.x, k_top5pc, 
                                                weight = n_responses)[1],
                        cor_k_top1pc = ~wtd.cor(.x, k_top1pc, 
                                                weight = n_responses)[1],
                        cor_k_rank_cond_parq1 = ~wtd.cor(.x, k_rank_cond_parq1, 
                                                         weight = n_responses)[1],
                        cor_k_rank_cond_parq2 = ~wtd.cor(.x, k_rank_cond_parq2, 
                                                         weight = n_responses)[1],
                        cor_k_rank_cond_parq3 = ~wtd.cor(.x, k_rank_cond_parq3, 
                                                         weight = n_responses)[1],
                        cor_k_rank_cond_parq4 = ~wtd.cor(.x, k_rank_cond_parq4, 
                                                         weight = n_responses)[1],
                        cor_k_rank_cond_parq5 = ~wtd.cor(.x, k_rank_cond_parq5, 
                                                         weight = n_responses)[1],
                        cor_kq5_cond_parq1 = ~wtd.cor(.x, kq5_cond_parq1, 
                                                      weight = n_responses)[1],
                        cor_kq5_cond_parq5 = ~wtd.cor(.x, kq5_cond_parq5, 
                                                      weight = n_responses)[1],
                        cor_par_q1 = ~wtd.cor(.x, par_q1, 
                                              weight = n_responses)[1],
                        cor_par_q5 = ~wtd.cor(.x, par_q5, 
                                              weight = n_responses)[1]),
                   .names = "{.col}_{.fn}") |>
  pivot_longer(names_to = "Variable", values_to = "Correlation", 1:3776) |>
  filter(str_detect(Variable, "mean")) |> 
  separate(Variable, c("Name", "Mobility"), sep = "_mean") |> 
  separate(Mobility, c("Drop", "Mobility"), sep = "or_") |> 
  select(-Drop) |> 
  pivot_wider(names_from = "Mobility", values_from = "Correlation") |> 
  select(-3)

css_correlations$Name <- gsub("(.+?)(\\..*)", "\\1", css_correlations$Name)

css_correlations <- left_join(varnames, css_correlations)

css_correlations <- css_correlations |> 
  select(-c("n_asked", "usable_m")) |> 
  filter(Name != "campus_id") |> 
  filter(Name != "n_responses")
```

```{r}
write.csv(css_correlations, "../data/css_correlations.csv", row.names = FALSE)
```

# CLEANUP TFS

library(haven)
#css <- read_spss("/Users/lawrence/Documents/heri_data/css/CSS.TRENDS.94.08.ARCHIVED DATA.SAV")


# heri_choice = 14235467
# restrict years = 4213528
# restrict css college = 1937983

## Demographics
heri_demographics <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/1 DEMOGRAPHICS.SAV")
heri_demographics <- heri_demographics |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()

## High school
heri_hs <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/2 HIGH SCHOOL.SAV")
heri_hs <- heri_hs |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()


## Choice
heri_choice <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/3 CHOICE.SAV")
heri_choice <- heri_choice |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()


## Plans
heri_plans <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/4 PLANS.SAV")
heri_plans <- heri_plans |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()


## View
heri_view <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/5 VIEW.SAV")
heri_view <- heri_view |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()


## Funds
heri_funds <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/6 FUNDS.SAV")
heri_funds <- heri_funds |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()


## Disagg
heri_disagg <- read_sav("/Users/lawrence/Documents/heri_data/tfs/master/7 DISAGG.SAV")
heri_disagg <- heri_disagg |> 
  filter(YEAR %in% 1995:2005) |> 
  filter(ACERECODE %in% acerecodes)

gc()



fwrite(heri_choice, "/users/lawrence/desktop/heri/tfs/choice.csv",
       row.names = FALSE)

fwrite(heri_demographics, "/users/lawrence/desktop/heri/tfs/demographics.csv",
       row.names = FALSE)

fwrite(heri_hs, "/users/lawrence/desktop/heri/tfs/hs.csv",
       row.names = FALSE)

fwrite(heri_plans, "/users/lawrence/desktop/heri/tfs/plans.csv",
       row.names = FALSE)

fwrite(heri_view, "/users/lawrence/desktop/heri/tfs/view.csv",
       row.names = FALSE)

fwrite(heri_funds, "/users/lawrence/desktop/heri/tfs/funds.csv",
       row.names = FALSE)

fwrite(heri_disagg, "/users/lawrence/desktop/heri/tfs/disagg.csv",
       row.names = FALSE)


## Constructs
FAC_INTERACTION
ACADEMIC_SELFCONCEPT
SOCIAL_SELFCONCEPT
SOCIAL_AGENCY
